---
title: "Parse Survey"
author: "Asa Bjorklund"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document:
    self_contained: true
    highlight: tango
    df_print: paged
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---


```{r}

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(xlsx))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(pheatmap))


```


```{r}
data = read.xlsx("ELIXIR SCO Training survey (réponses).xlsx", sheetIndex = 1, header = F)
colnames = as.character(data[1,])
data = data[-1,]

```


# Responses per country

```{r}

data$Country = data[,4]
data$Country = sub("ELIXIR-","", data$Country)
data$Country = sub("FRANCE","France", data$Country)
data$Country = sub("The ","", data$Country)
data$Country[grep("UK", data$Country)] = "United Kingdom"
data$Country[grep("German", data$Country)] = "Germany"
data$Country[grep("United States", data$Country)] = "USA"
data$Country[grep("INB Spain", data$Country)] = "Spain"

t = table(data$Country)
data = data[order(data$Country),]
data$nCountry = paste(data$Country, unlist(sapply(t, function(x) 1:x)), sep = "-")

par(mar=c(c(10,2,2,2)))
barplot(table(data$Country), las = 2, main = "Responses per country")

all.countries = unique(data$Country)
```

Responses with no courses:

```{r}
a = which(is.na(data[,5]))

b = grep("None", data[,5])

print(data$Country[c(a,b)])
```



# Parse per course


For each course, parse the info in all parts of the questions, also for "Rising need" or "I can teach"

```{r}


parse_course_url = function(x){
  s = strsplit(x,"\n")[[1]]
  s = gsub("Course *#[0-9] *[:-] *", "",s)
  s = s[nchar(s)>1]
  return(s)
}


options = c("Course #1", "Course #2", "Course #3", "Course #4", "Rising need", "I can teach")
#parse sections 2-11, skipping 7
section.idx = c(2:6,8:11)
section.trans = c("Teacher","Duration","nStudents", "SCO", "SRT","EntryReq","Strategies","Compute", "Topics")
names(section.trans)=section.idx

option = sapply(colnames, str_extract, "\\[.*\\]")
option = gsub("\\]", "", option)
option = gsub("\\[", "", option)
sec = sapply(colnames, str_extract ,"^\\d+")
new.name = paste(section.trans[sec],option, sep = " : ")
new.name[is.na(option)] = colnames[is.na(option)]
colnames(data) = new.name

parse_line = function(x){
  urls = parse_course_url(x[,5])
  datasets = parse_course_url(x[,37])
  email = x[,2]
  s = sapply(options, grepl, x)
  s = apply(s,2, function(y) new.name[y])
  s = lapply(s, function(y) y[!grepl("^\\d",y)]) # remove names starting with number
  
  courses = s[grepl("Course", names(s))]  
  courses = courses[sapply(courses,length)>0]
  nC = length(courses)
  if(nC == 0){
    return(list(country=x[,85],name=x[,86], needs=s$`Rising need`, urls = urls, courses = NULL, email = email))
  }
  
  names(courses) = paste(x[,86], names(courses), sep = ":")
  courses = lapply(courses, function(y) c(y, paste0("Country : ",x[,85])))
  courses = lapply(courses, function(y) c(y, paste0("Email : ",x[,2])))
  if (length(urls) != length(courses)){
    print("Different number of urls. ")
    print(email)
    print(urls)
    print(names(courses))
  }
  if (length(datasets) != length(courses)){
    print("Different number of datasets ")
    print(email)
    print(datasets)
    print(names(courses))
  }
  for (i in 1:length(courses)){
    courses[[i]] = c(courses[[i]], paste0("Url : ", urls[min(i, length(urls))]))
    courses[[i]] = c(courses[[i]], paste0("Dataset : ", datasets[min(i, length(datasets))]))
  }
  
  return(list(country=x[,85],name=x[,86], needs=s$`Rising need`, urls = urls, courses = courses))
   
}


all.data = list()
for (i in 1:nrow(data)){
  all.data[[i]] = parse_line(data[i,])
}
```

OBS! USA-1 - same course for 2019,20,21,22 - only one url.


Comment:
All other courses on single cell I know are on glittr.org: https://glittr.org/?search=single+cell&per_page=25&sort_by=stargazers&sort_direction=desc

Multiple galaxyproject entries from Netherlands possibly pointing to the same course.


```{r, fig.width=12, fig.height=10}
courses = sapply(all.data, function(x) x$courses)
courses = unlist(courses, recursive = F)

all.opts = new.name[!grepl("^\\d", new.name)][-c(1:4)]
all.opts = all.opts[-75]

stats = Reduce(rbind,lapply(courses, function(x) all.opts %in% x))
colnames(stats) = all.opts
rownames(stats) = names(courses)
stats = stats + 0
stats = t(stats)

annot = data.frame(country = sapply(strsplit(names(courses),"-"), function(x) x[1]))
rownames(annot) = names(courses)
```


#### Topics

```{r, fig.width=12, fig.height=10}
sel.topics = c(grep("SCO", rownames(stats)), grep("SRT", rownames(stats)), grep("Topics", rownames(stats)))



#pheatmap(stats[sel.topics,], cluster_rows = F, fontsize = 6, annotation_col = annot)
```

One of the Dutch courses is the resource page https://www.singlecell.nl/home, it has all topics included in the list. Remove from summary stats.


Grouped by country

```{r}
course2country = lapply(courses, function(x) gsub("Country : ", "", x[grepl("Country",x)]))

tmp = data.frame(t(stats))
tmp$Country = course2country

tmp = tmp[rownames(tmp) != "Netherlands-5:Course #2",]

tmp2 = tmp %>% group_by(Country) %>% summarise_each(list(sum))

stats.country = data.frame(tmp2)
rownames(stats.country) = tmp2$Country
stats.country = stats.country[,-1]
stats.country = t((stats.country>0) + 0)
```


```{r}
#pheatmap(stats.country[sel.topics,], cluster_rows = F, fontsize = 6)
```


#### Teaching methods

```{r, fig.width=12, fig.height=10}
sel.teaching = c(grep("Strategies", rownames(stats)), grep("Compute", rownames(stats)))



#pheatmap(stats[sel.teaching,], cluster_rows = F, fontsize = 6, annotation_col = annot)
```

Grouped by country

```{r}
#pheatmap(stats.country[sel.teaching,], cluster_rows = F, fontsize = 6)
```

## Rising need

Per country

```{r}
needs = lapply(all.data, function(x) x$needs)
countries = unlist(lapply(all.data, function(x) x$country))

s = split(needs, countries)
s = lapply(s, unlist)

all.cat = rownames(stats)[c(grep("SCO", rownames(stats)), grep("SRT", rownames(stats)))]

need.stats = Reduce(cbind, lapply(s, function(x) all.cat %in% x)) + 0
rownames(need.stats) = all.cat
colnames(need.stats) = names(s)

pheatmap(need.stats, cluster_rows = F)
```


#### Print to csv

```{r}
all.courses = unlist(lapply(all.data, function(x) x$courses), recursive = F)

course.names = colnames(stats)

sel.columns = c("Country","Email","Url","Dataset")
pattern = paste(sel.columns, collapse = "|")

extra = Reduce(rbind, lapply(all.courses, function(x) x[grepl(pattern,x)]))
colnames(extra) = sel.columns


out = data.frame(cbind(extra),t(stats))
rownames(out) = colnames(stats)
write.csv(out, "parsed_courses.csv")
```

# Challenges

Summarized per inidivual response

## Pedagogical

```{r}
tmp = data.frame(sort(table(unlist(strsplit(data[,81],", ")))))

tmp = tmp[!grepl("especially",tmp$Var1),]

ggplot(tmp, aes(x=Var1, y=Freq)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic()
```

## Technical

```{r}
tmp = data.frame(sort(table(unlist(strsplit(data[,82],"Issues | Data")))), stringsAsFactors = F)
# remove empty
tmp = tmp[nchar(as.character(tmp$Var1))>0,]


#add back the "Issues"
starts = c("on", "in", "due", "with")
is.issue = rowSums(sapply(starts, function(x) startsWith(as.character(tmp$Var1),x)))
tmp$Var1 = as.character(tmp$Var1)
tmp$Var1[is.issue>0] = paste0("Issues ",as.character(tmp$Var1[is.issue>0]))
# max length 80
tmp$Var1 = substr(tmp$Var1,1,80)
tmp$Var1 = factor(tmp$Var1, levels = tmp$Var1) 
ggplot(tmp, aes(x=Var1, y=Freq)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic()
```

# Replot after annotation

Remove all duplicate courses, and websites etc that are not really courses, and use the country of origin for the course.

```{r}
courses2 = read.csv("Training_survey_parsed_courses - parsed_courses.csv", header = T, row.names = 1)

# remove duplicates and non-courses
courses2 = courses2[nchar(courses2$Is.duplicate) ==0,]
courses2 = courses2[nchar(courses2$Not.a.course) ==0,]


```

## Courses per country

```{r}
par(mar=c(10,2,2,2))
t = table(courses2$Country.Edited)
barplot(t, las=2, main = "Courses given per country")
```

## Topics / Technologies

### By course

```{r, fig.width=12, fig.height=12}
sel.topics = c(grep("SCO", colnames(courses2)), grep("SRT", colnames(courses2)), grep("Topics", colnames(courses2)))

annot = data.frame(country = courses2$Country.Edited)
annot$country = gsub("Country : ","", annot$country)
rownames(annot) = rownames(courses2)


pheatmap(t(courses2[,sel.topics]), cluster_rows = F, cluster_cols = T, fontsize = 6, annotation_col = annot) 

```

Most/least covered topics

```{r, fig.width=12, fig.height=12}

tmp = data.frame(count = sort(colSums(courses2[,sel.topics])))
tmp$topic = factor(rownames(tmp),levels = rownames(tmp))
tmp$group = unlist(lapply(strsplit(rownames(tmp),"\\.."), function(x) x[1]))
ggplot(tmp, aes(x=topic,y=count, fill=group)) + geom_bar(stat="identity", width=0.5) + coord_flip() 
```


### By country

```{r, fig.width=12, fig.height=12}
tmp = courses2[,sel.topics]
tmp$country = courses2$Country.Edited
tmp2 = tmp %>% group_by(country) %>% summarise_each(list(sum))
tmp2 = data.frame(tmp2)
rownames(tmp2) = tmp2$country
tmp2 = tmp2[,-1]

pheatmap(t(tmp2>0)+1, cluster_rows = F, fontsize = 6)

```

# Teaching methods

```{r, fig.width=12, fig.height=12}
sel.teaching = c(grep("Strategies", colnames(courses2)), grep("Compute", colnames(courses2)))

tmp = data.frame(count = sort(colSums(courses2[,sel.teaching])))
tmp$group = unlist(lapply(strsplit(rownames(tmp),"\\.."), function(x) x[1]))
tmp = tmp[!grepl("Other",rownames(tmp)),]

tmp$topic = gsub("Strategies...","",rownames(tmp))
tmp$topic = gsub("Compute...","",tmp$topic)
tmp$topic = factor(tmp$topic,levels = tmp$topic)

ggplot(tmp[tmp$group == "Compute",], aes(x=topic,y=count, fill=group)) + geom_bar(stat="identity", width=0.5) + coord_flip()  + theme_classic() + ggtitle("Compute")
ggplot(tmp[tmp$group == "Strategies",], aes(x=topic,y=count, fill=group)) + geom_bar(stat="identity", width=0.5) + coord_flip()  + theme_classic() + ggtitle("Teaching strategies")
```

```{r}
tmp = courses2[,sel.teaching]
tmp$country = courses2$Country.Edited
tmp2 = tmp %>% group_by(country) %>% summarise_each(list(sum))
tmp2 = data.frame(tmp2)
rownames(tmp2) = tmp2$country
tmp2 = tmp2[,-1]

pheatmap(t(tmp2>0)+1, cluster_rows = F, fontsize = 6)
```



# Course duration

```{r}
sel.duration = grep("Duration", colnames(courses2))


tmp = data.frame(count = colSums(courses2[,sel.duration]))
rownames(tmp) = sub("Duration...","",rownames(tmp))
tmp$topic = factor(rownames(tmp),levels = rownames(tmp))
tmp$group = c(rep("Days",5),rep("Consecutive",2),rep("Yearly",3))
ggplot(tmp, aes(x=topic,y=count, fill = group)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic()
```



# Participants

```{r}
sel.nS = grep("nS", colnames(courses2))


tmp = data.frame(count = colSums(courses2[,sel.nS]))
rownames(tmp) = sub("nStudents...","",rownames(tmp))
tmp$topic = factor(rownames(tmp),levels = rownames(tmp))
ggplot(tmp, aes(x=topic,y=count)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic()
```


# Entry requirements

```{r}
sel.nS = grep("EntryReq", colnames(courses2))

tmp = data.frame(count = colSums(courses2[,sel.nS]))
rownames(tmp) = sub("EntryReq...","",rownames(tmp))
tmp$topic = factor(rownames(tmp),levels = rownames(tmp))
ggplot(tmp, aes(x=topic,y=count)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic()
```

```{r}
sessionInfo()
```

