---
title: "Parse Survey"
author: "Asa Bjorklund"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document:
    self_contained: true
    highlight: tango
    df_print: paged
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---


```{r}

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(xlsx))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(pheatmap))


```


```{r}
data = read.xlsx("ELIXIR SCO Training survey (réponses).xlsx", sheetIndex = 1, header = F)
colnames = as.character(data[1,])
data = data[-1,]

```


# Responses per country

```{r}

data$Country = data[,4]
data$Country = sub("ELIXIR-","", data$Country)
data$Country = sub("FRANCE","France", data$Country)
data$Country = sub("The ","", data$Country)
data$Country[grep("UK", data$Country)] = "United Kingdom"
data$Country[grep("German", data$Country)] = "Germany"
data$Country[grep("United States", data$Country)] = "USA"
data$Country[grep("INB Spain", data$Country)] = "Spain"

t = table(data$Country)
data = data[order(data$Country),]
data$nCountry = paste(data$Country, unlist(sapply(t, function(x) 1:x)), sep = "-")


barplot(table(data$Country), las = 2, main = "Responses per country")
```

Responses with no courses:

```{r}
a = which(is.na(data[,5]))

b = grep("None", data[,5])

print(data$Country[c(a,b)])
```



# Responses per course


For each course, parse the info in all parts of the questions, also for "Rising need" or "I can teach"

```{r}


parse_course_url = function(x){
  s = strsplit(x,"\n")[[1]]
  s = gsub("Course *#[0-9] *[:-] *", "",s)
  s = s[nchar(s)>1]
  return(s)
}
course_list = sapply(data[,5], parse_course_url)


options = c("Course #1", "Course #2", "Course #3", "Course #4", "Rising need", "I can teach")
#parse sections 2-11, skipping 7
section.idx = c(2:6,8:11)
section.trans = c("Teacher","Duration","nStudents", "SCO", "SRT","EntryReq","Strategies","Compute", "Topics")
names(section.trans)=section.idx

option = sapply(colnames, str_extract, "\\[.*\\]")
option = gsub("\\]", "", option)
option = gsub("\\[", "", option)
sec = sapply(colnames, str_extract ,"^\\d+")
new.name = paste(section.trans[sec],option, sep = " : ")
new.name[is.na(option)] = colnames[is.na(option)]
colnames(data) = new.name

parse_line = function(x){
  urls = parse_course_url(x[,5])
  email = x[,2]
  s = sapply(options, grepl, x)
  s = apply(s,2, function(y) new.name[y])
  s = lapply(s, function(y) y[!grepl("^\\d",y)]) # remove names starting with number
  
  courses = s[grepl("Course", names(s))]  
  courses = courses[sapply(courses,length)>0]
  nC = length(courses)
  if(nC == 0){
    return(list(country=x[,85],name=x[,86], needs=s$`Rising need`, urls = urls, courses = NULL, email = email))
  }
  
  names(courses) = paste(x[,86], names(courses), sep = ":")
  courses = lapply(courses, function(y) c(y, paste0("Country : ",x[,85])))
  courses = lapply(courses, function(y) c(y, paste0("Email : ",x[,2])))
  if (length(urls) != length(courses)){
    print("Different number of urls. ")
    print(email)
    print(urls)
    print(names(courses))
  }
  for (i in 1:length(courses)){
    courses[[i]] = c(courses[[i]], paste0("Url : ", urls[min(i, length(urls))]))
  }
  
  
  return(list(country=x[,85],name=x[,86], needs=s$`Rising need`, urls = urls, courses = courses))
   
}



all.data = list()
for (i in 1:nrow(data)){
  all.data[[i]] = parse_line(data[i,])
}
```

OBS! USA-1 - same course for 2019,20,21,22 - only one url.


Comment:
All other courses on single cell I know are on glittr.org: https://glittr.org/?search=single+cell&per_page=25&sort_by=stargazers&sort_direction=desc

Multiple galaxyproject entries from Netherlands possibly pointing to the same course.


```{r, fig.width=12, fig.height=10}
courses = sapply(all.data, function(x) x$courses)
courses = unlist(courses, recursive = F)

all.opts = new.name[!grepl("^\\d", new.name)][-c(1:4)]
all.opts = all.opts[-75]

stats = Reduce(rbind,lapply(courses, function(x) all.opts %in% x))
colnames(stats) = all.opts
rownames(stats) = names(courses)
stats = stats + 0
stats = t(stats)

annot = data.frame(country = sapply(strsplit(names(courses),"-"), function(x) x[1]))
rownames(annot) = names(courses)
```


# Topics

```{r, fig.width=12, fig.height=10}
sel.topics = c(grep("SCO", rownames(stats)), grep("SRT", rownames(stats)), grep("Topics", rownames(stats)))



pheatmap(stats[sel.topics,], cluster_rows = F, fontsize = 6, annotation_col = annot)
```

One of the Dutch courses is the resource page https://www.singlecell.nl/home, it has all topics included in the list. Remove from summary stats.


Grouped by country

```{r}
course2country = lapply(courses, function(x) gsub("Country : ", "", x[grepl("Country",x)]))

tmp = data.frame(t(stats))
tmp$Country = course2country

tmp = tmp[rownames(tmp) != "Netherlands-5:Course #2",]

tmp2 = tmp %>% group_by(Country) %>% summarise_each(list(sum))

stats.country = data.frame(tmp2)
rownames(stats.country) = tmp2$Country
stats.country = stats.country[,-1]
stats.country = t((stats.country>0) + 0)
```


```{r}
pheatmap(stats.country[sel.topics,], cluster_rows = F, fontsize = 6)
```


# Teaching methods

```{r, fig.width=12, fig.height=10}
sel.teaching = c(grep("Strategies", rownames(stats)), grep("Compute", rownames(stats)))



pheatmap(stats[sel.teaching,], cluster_rows = F, fontsize = 6, annotation_col = annot)
```

Grouped by country

```{r}
pheatmap(stats.country[sel.teaching,], cluster_rows = F, fontsize = 6)
```

# Rising need

Per country

```{r}
needs = lapply(all.data, function(x) x$needs)
countries = unlist(lapply(all.data, function(x) x$country))

s = split(needs, countries)
s = lapply(s, unlist)

all.cat = rownames(stats)[c(grep("SCO", rownames(stats)), grep("SRT", rownames(stats)))]

need.stats = Reduce(cbind, lapply(s, function(x) all.cat %in% x)) + 0
rownames(need.stats) = all.cat
colnames(need.stats) = names(s)

pheatmap(need.stats, cluster_rows = F)
```


# Print courses to a csv.

```{r}
all.courses = unlist(lapply(all.data, function(x) x$courses), recursive = F)

all.courses = lapply(all.courses, function(x) Reduce(rbind, strsplit(x, " : ")))
headers = unique(unlist(lapply(all.courses, function(x) x[,1])))


```


```{r}
sessionInfo()
```

