---
title: "Training Survey Overview"
author: "Asa Bjorklund"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document:
    self_contained: true
    highlight: tango
    df_print: paged
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

```{r stp1, echo=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, fig.align='center', fig.path='figure/', progress=FALSE, verbose=FALSE, message = FALSE, warning = FALSE, error = FALSE, fig.width = 8, fig.height = 5.5)
```



```{r, echo=FALSE}

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(xlsx))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(pheatmap))


```


```{r, echo=FALSE}
data = read.xlsx("ELIXIR SCO Training survey (réponses).xlsx", sheetIndex = 1, header = F)
colnames = as.character(data[1,])
data = data[-1,]

```


# Summary by repsponses

### Responses per country

```{r, echo=FALSE}

data$Country = data[,4]
data$Country = sub("ELIXIR-","", data$Country)
data$Country = sub("FRANCE","France", data$Country)
data$Country = sub("The ","", data$Country)
data$Country[grep("UK", data$Country)] = "United Kingdom"
data$Country[grep("German", data$Country)] = "Germany"
data$Country[grep("United States", data$Country)] = "USA"
data$Country[grep("INB Spain", data$Country)] = "Spain"

t = table(data$Country)
data = data[order(data$Country),]
data$nCountry = paste(data$Country, unlist(sapply(t, function(x) 1:x)), sep = "-")

tmp = data.frame(table(data$Country))
colnames(tmp) = c("Country","Counts")
ggplot(tmp, aes(x=Country,y=Counts)) + geom_bar(stat="identity", width=0.5, fill = "blue") + theme_classic() + coord_flip()

all.countries = unique(data$Country)
```



```{r, echo=FALSE, results='hide'}
# Responses with no country
a = which(is.na(data[,5]))

b = grep("None", data[,5])

cat("No courses listed for ", paste(data$Country[c(a,b)], collapse = ", "))
```



### Pedagogical Challenges


```{r, echo=FALSE}
# remove commas in the answers to split on ,
tmp1 = data[,81]
tmp1 = gsub("interactive,","interactive;",tmp1)

tmp = data.frame(sort(table(unlist(strsplit(tmp1,", ")))))
colnames(tmp) = c("Response","Counts")

# put back comma
tmp$Response = gsub(";",",", tmp$Response)

# only one "other"
other = tmp$Response[grepl("Not sure yet", tmp$Response)]
tmp$Response[grepl("Not sure yet", tmp$Response)] = "Other"

# add in linebreaks
tmp$Response = sapply(tmp$Response, function(x) paste(strwrap(x,50), collapse="\n"))

# order by size
lev = tmp$Response[!grepl("Other", tmp$Response)][order(tmp$Count[!grepl("Other", tmp$Response)], decreasing = F)]
tmp$Response = factor(tmp$Response, levels = c("Other", lev))


colnames(tmp) = c("Response","Counts")
ggplot(tmp, aes(x=Response, y=Counts)) + geom_bar(stat="identity", width=0.5, fill = "blue") + coord_flip() + theme_classic() + theme(axis.text.y = element_text(size = 8))
```

Other includes answer `r other`.

### Technical Challenges

```{r, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
#tmp = data.frame(sort(table(unlist(strsplit(data[,82],"Issues | Data")))), stringsAsFactors = F)
tmp = data.frame(sort(table(unlist(strsplit(data[,82],"Issues ")))), stringsAsFactors = F)

# remove empty
tmp = tmp[nchar(as.character(tmp$Var1))>0,]
colnames(tmp) = c("Response","Counts")

#add back the "Issues"
starts = c("on", "in", "due", "with")
is.issue = rowSums(sapply(starts, function(x) startsWith(as.character(tmp$Response),x)))
tmp$Response = as.character(tmp$Response)
tmp$Response[is.issue>0] = paste0("Issues ",as.character(tmp$Response[is.issue>0]))

other = tmp$Response[is.issue == 0]
tmp$Response[is.issue == 0] = "Other"

# add in linebreaks
tmp$Response = sapply(tmp$Response, function(x) paste(strwrap(x,60), collapse="\n"))

# order by size
lev = tmp$Response[!grepl("Other", tmp$Response)][order(tmp$Count[!grepl("Other", tmp$Response)], decreasing = F)]
tmp$Response = factor(tmp$Response, levels = c("Other", lev))

ggplot(tmp, aes(x=Response, y=Counts)) + geom_bar(stat="identity", width=0.5, fill = "blue") + coord_flip() + theme_classic()
```





```{r, echo=FALSE, results='hide'}
# Parse information by course

#Instead of listing results by the respondents, extract the information from each individual course. 

#For each course, parse the info in all parts of the questions, also for "Rising need" or "I can teach"

parse_course_url = function(x){
  s = strsplit(x,"\n")[[1]]
  s = gsub("Course *#[0-9] *[:-] *", "",s)
  s = s[nchar(s)>1]
  return(s)
}


options = c("Course #1", "Course #2", "Course #3", "Course #4", "Rising need", "I can teach")
#parse sections 2-11, skipping 7
section.idx = c(2:6,8:11)
section.trans = c("Teacher","Duration","nStudents", "SCO", "SRT","EntryReq","Strategies","Compute", "Topics")
names(section.trans)=section.idx

option = sapply(colnames, str_extract, "\\[.*\\]")
option = gsub("\\]", "", option)
option = gsub("\\[", "", option)
sec = sapply(colnames, str_extract ,"^\\d+")
new.name = paste(section.trans[sec],option, sep = " : ")
new.name[is.na(option)] = colnames[is.na(option)]
colnames(data) = new.name

parse_line = function(x){
  urls = parse_course_url(x[,5])
  datasets = parse_course_url(x[,37])
  email = x[,2]
  s = sapply(options, grepl, x)
  s = apply(s,2, function(y) new.name[y])
  s = lapply(s, function(y) y[!grepl("^\\d",y)]) # remove names starting with number
  
  courses = s[grepl("Course", names(s))]  
  courses = courses[sapply(courses,length)>0]
  nC = length(courses)
  if(nC == 0){
    return(list(country=x[,85],name=x[,86], needs=s$`Rising need`, urls = urls, courses = NULL, email = email))
  }
  
  names(courses) = paste(x[,86], names(courses), sep = ":")
  courses = lapply(courses, function(y) c(y, paste0("Country : ",x[,85])))
  courses = lapply(courses, function(y) c(y, paste0("Email : ",x[,2])))
  if (length(urls) != length(courses)){
    print("Different number of urls. ")
    print(email)
    print(urls)
    print(names(courses))
  }
  if (length(datasets) != length(courses)){
    print("Different number of datasets ")
    print(email)
    print(datasets)
    print(names(courses))
  }
  for (i in 1:length(courses)){
    courses[[i]] = c(courses[[i]], paste0("Url : ", urls[min(i, length(urls))]))
    courses[[i]] = c(courses[[i]], paste0("Dataset : ", datasets[min(i, length(datasets))]))
  }
  
  return(list(country=x[,85],name=x[,86], needs=s$`Rising need`, urls = urls, courses = courses))
   
}


all.data = list()
for (i in 1:nrow(data)){
  all.data[[i]] = parse_line(data[i,])
}
```

```{r, echo=FALSE, results='hide'}
#OBS! USA-1 - same course for 2019,20,21,22 - only one url.


#Comment:
#All other courses on single cell I know are on glittr.org: https://glittr.org/?search=single+cell&per_page=25&sort_by=stargazers&sort_direction=desc

#Multiple galaxyproject entries from Netherlands possibly pointing to the same course.

# One of the Dutch courses is the resource page https://www.singlecell.nl/home, it has all topics included in the list. Remove from summary stats.
```

```{r,  echo=FALSE}
courses = sapply(all.data, function(x) x$courses)
courses = unlist(courses, recursive = F)

all.opts = new.name[!grepl("^\\d", new.name)][-c(1:4)]
all.opts = all.opts[-75]

stats = Reduce(rbind,lapply(courses, function(x) all.opts %in% x))
colnames(stats) = all.opts
rownames(stats) = names(courses)
stats = stats + 0
stats = t(stats)

annot = data.frame(country = sapply(strsplit(names(courses),"-"), function(x) x[1]))
rownames(annot) = names(courses)
```


```{r, echo=FALSE}
# Topics

sel.topics = c(grep("SCO", rownames(stats)), grep("SRT", rownames(stats)), grep("Topics", rownames(stats)))



#pheatmap(stats[sel.topics,], cluster_rows = F, fontsize = 6, annotation_col = annot)
```


```{r, echo=FALSE}
# Grouped by country
course2country = lapply(courses, function(x) gsub("Country : ", "", x[grepl("Country",x)]))

tmp = data.frame(t(stats))
tmp$Country = course2country

tmp = tmp[rownames(tmp) != "Netherlands-5:Course #2",]

tmp2 = tmp %>% group_by(Country) %>% summarise_each(list(sum))

stats.country = data.frame(tmp2)
rownames(stats.country) = tmp2$Country
stats.country = stats.country[,-1]
stats.country = t((stats.country>0) + 0)

#pheatmap(stats.country[sel.topics,], cluster_rows = F, fontsize = 6)
```


```{r,  echo=FALSE}
#### Teaching methods
sel.teaching = c(grep("Strategies", rownames(stats)), grep("Compute", rownames(stats)))

#pheatmap(stats[sel.teaching,], cluster_rows = F, fontsize = 6, annotation_col = annot)

# Grouped by country
#pheatmap(stats.country[sel.teaching,], cluster_rows = F, fontsize = 6)
```

### Technologies with a rising need

Areas where the respondents sees a rising need, grouped by country

SCO = Single Cell Omics, SRT = Spatially Resolved Transcriptomics

```{r, echo=FALSE, fig.height=5}
needs = lapply(all.data, function(x) x$needs)
countries = unlist(lapply(all.data, function(x) x$country))

s = split(needs, countries)
s = lapply(s, unlist)

all.cat = rownames(stats)[c(grep("SCO", rownames(stats)), grep("SRT", rownames(stats)))]

need.stats = Reduce(cbind, lapply(s, function(x) all.cat %in% x)) + 0
rownames(need.stats) = all.cat
colnames(need.stats) = names(s)

pheatmap(need.stats, cluster_rows = F, legend_breaks = c(0,1), legend_labels = c("no","yes"))
```



```{r, echo=FALSE}
#### Print to csv
all.courses = unlist(lapply(all.data, function(x) x$courses), recursive = F)

course.names = colnames(stats)

sel.columns = c("Country","Email","Url","Dataset")
pattern = paste(sel.columns, collapse = "|")

extra = Reduce(rbind, lapply(all.courses, function(x) x[grepl(pattern,x)]))
colnames(extra) = sel.columns


out = data.frame(cbind(extra),t(stats))
rownames(out) = colnames(stats)
write.csv(out, "parsed_courses.csv")
```

# Summary by course

Instead of listing results by the respondents, extract the information from each individual course. Each response contains up to 4 courses.

Manual filtering of duplicate course instances, correct country of origin and examples of entries that are not really courses.



```{r, echo=FALSE}
courses2 = read.table("Training_survey_parsed_courses - parsed_courses.tsv", header = T, row.names = 1, sep = "\t", quote = "", comment.char = "")

# remove duplicates and non-courses
courses2 = courses2[nchar(courses2$Is.duplicate) ==0,]
courses2 = courses2[nchar(courses2$Not.a.course) ==0,]

# revert back to the old colnames in stats, not with ...
old.names = rownames(stats)
colnames(courses2) = c(colnames(courses2)[1:13],old.names)

```

In total `r nrow(courses2)` courses are included.

### Courses given per country

```{r, echo=FALSE}

tmp = data.frame(table(courses2$`Country.Edited`))
colnames(tmp) = c("Country","Counts")

ggplot(tmp, aes(x=Country,y=Counts)) + geom_bar(stat="identity", width=0.5, fill = "blue") + theme_classic() + coord_flip() + theme(axis.title.y = element_blank())

```


### Online material

```{r, fig.width=6, fig.height=6}
tmp = data.frame(Online_material = grepl("Yes", courses2$Online.material), Online_lectures = grepl("Yes", courses2$Online.lectures))
tmp$Country = courses2$Country.Edited
library(grid)

#table(tmp[,1:2])
setHook("grid.newpage", function() pushViewport(viewport(x=1,y=1,width=0.9, height=0.9, name="vp", just=c("right","top"))), action="prepend")
pheatmap(table(tmp[,1:2]), display_numbers = T, cluster_rows = F, cluster_cols = F, number_format = "%d", fontsize_number = 20, legend = F)

setHook("grid.newpage", NULL, "replace")
grid.text("Online Lectures", y=-0.07, gp=gpar(fontsize=16))
grid.text("Online Course Material", x=-0.07, rot=90, gp=gpar(fontsize=16))


```

Per country/course

```{r, fig.width=6, fig.height=8}
rownames(tmp) = rownames(courses2)
annot2 = data.frame(Country = tmp$Country)
pheatmap(tmp[,1:2]+0, cluster_cols = F, annotation_row = annot, treeheight_row = 0, legend_breaks = c(0,1), legend_labels = c("no","yes"))
```




### Topics / Technologies - By course



```{r}
sel.topics = c(grep("SCO", colnames(courses2)), grep("SRT", colnames(courses2)), grep("Topics", colnames(courses2)))

annot = data.frame(country = courses2$`Country.Edited`)
annot$country = gsub("Country : ","", annot$country)
rownames(annot) = rownames(courses2)

```

```{r, echo=FALSE}
pheatmap(t(courses2[,sel.topics]), cluster_rows = F, cluster_cols = T, fontsize = 6, annotation_col = annot, legend_breaks = c(0,1), legend_labels = c("no","yes"), show_colnames = F)

```


### Topics / Technologies - Number of courses 

```{r}

tmp = data.frame(Count = sort(colSums(courses2[,sel.topics])))
tmp$Topic = factor(rownames(tmp),levels = rownames(tmp))
tmp$group = unlist(lapply(strsplit(rownames(tmp)," : "), function(x) x[1]))
ggplot(tmp, aes(x=Topic,y=Count, fill=group)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic() + theme(axis.title.y = element_blank())
```

Also as two separate plots, one for Topics and one for Technologies

```{r}
ggplot(tmp[tmp$group != "Topics",], aes(x=Topic,y=Count, fill=group)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic() + theme(axis.title.y = element_blank())

ggplot(tmp[tmp$group == "Topics",], aes(x=Topic,y=Count)) + geom_bar(stat="identity", width=0.5, fill = "blue") + coord_flip() + theme_classic() + theme(axis.title.y = element_blank())
```



### Topics / Technologies - By country

```{r}
tmp = courses2[,sel.topics]
tmp$country = courses2$`Country.Edited`
tmp2 = tmp %>% group_by(country) %>% summarise_each(list(sum))
tmp2 = data.frame(tmp2)
rownames(tmp2) = tmp2$country
tmp2 = tmp2[,-1]

pheatmap(t(tmp2>0)+0, cluster_rows = F, fontsize = 6, legend_breaks = c(0,1), legend_labels = c("no","yes"))

```

### Teaching methods - Computational

```{r}
sel.teaching = c(grep("Strategies", colnames(courses2)), grep("Compute", colnames(courses2)))

tmp = data.frame(Count = sort(colSums(courses2[,sel.teaching])))
tmp$Group = unlist(lapply(strsplit(rownames(tmp)," : "), function(x) x[1]))
tmp = tmp[!grepl("Other",rownames(tmp)),]

tmp$Topic = gsub("Strategies : ","",rownames(tmp))
tmp$Topic = gsub("Compute : ","",tmp$Topic)
tmp$Topic = factor(tmp$Topic,levels = tmp$Topic)

ggplot(tmp[tmp$Group == "Compute",], aes(x=Topic,y=Count)) + geom_bar(stat="identity", width=0.5, fill = "red") + coord_flip()  + theme_classic()  + theme(axis.title.y = element_blank())
```

### Teaching methods - Strategies

```{r}
ggplot(tmp[tmp$Group == "Strategies",], aes(x=Topic,y=Count)) + geom_bar(stat="identity", width=0.5, fill = "red") + coord_flip()  + theme_classic() + theme(axis.title.y = element_blank())
```

```{r, echo=FALSE, results='hide'}
tmp = courses2[,sel.teaching]
tmp$country = courses2$Country.Edited
tmp2 = tmp %>% group_by(country) %>% summarise_each(list(sum))
tmp2 = data.frame(tmp2)
rownames(tmp2) = tmp2$country
tmp2 = tmp2[,-1]

#pheatmap(t(tmp2>0)+1, cluster_rows = F, fontsize = 6, legend = FALSE)
```



### Course duration

```{r, echo=FALSE}
sel.duration = grep("Duration", colnames(courses2))


tmp = data.frame(Count = colSums(courses2[,sel.duration]))
rownames(tmp) = sub("Duration : ","",rownames(tmp))
tmp$Topic = factor(rownames(tmp),levels = rownames(tmp))
tmp$Group = c(rep("Course days",5),rep("Consecutive",2),rep("Times per year",3))
ggplot(tmp, aes(x=Topic,y=Count, fill = Group)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_classic() + theme(axis.title.y = element_blank())
```



### Participants

```{r, echo=FALSE}
sel.nS = grep("nS", colnames(courses2))


tmp = data.frame(count = colSums(courses2[,sel.nS]))
rownames(tmp) = sub("nStudents...","",rownames(tmp))
tmp$topic = factor(rownames(tmp),levels = rownames(tmp))
ggplot(tmp, aes(x=topic,y=count)) + geom_bar(stat="identity", width=0.5, fill = "blue") + coord_flip() + theme_classic() + theme(axis.title.y = element_blank())
```


### Entry requirements

```{r, echo=FALSE}
sel.nS = grep("EntryReq", colnames(courses2))

tmp = data.frame(count = sort(colSums(courses2[,sel.nS])))
rownames(tmp) = sub("EntryReq...","",rownames(tmp))
tmp$topic = factor(rownames(tmp),levels = rownames(tmp))
ggplot(tmp, aes(x=topic,y=count)) + geom_bar(stat="identity", width=0.5, fill = "blue") + coord_flip() + theme_classic() + theme(axis.title.y = element_blank())
```



# Conclusions


* Most courses are given once a year, for 1 week with 20-40 students.
* Biggest challenges are non-uniform audiences and different operating systems
* Mainly R based courses.
* Few courses covering other SC-omics than RNA-seq, like ATAC, CITE, VDJ etc.
* For spatial data, mainly visium covered in courses.




```{r, echo=FALSE, results='hide'}
sessionInfo()
```

